---
title: "Hw05ST430Yu"
author: "Haozhe (Jerry) Yu"
date: "`r Sys.Date()`"
output: 
  pdf_document:
    latex_engine: xelatex
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggpmisc)
library(dplyr)
library(tidyverse)
library(tinytex)
library(usethis)
library(lmtest)
library(car)
```

# Question 1

```{r import 1}
software <- as_tibble(read_table("Datasets/software.txt",
            col_names = TRUE,
            ))
```

## 1A. Fit a model in which sales last quarter is ignored. We want to know whether software package has any effect on sales.

```{r SLR}
softwareslr <- lm(SalesThisQuarter~as.factor(Software),software)

summary(softwareslr)

anova(softwareslr)

str(summary(softwareslr))

summary(softwareslr)$coefficients

summary(softwareslr)[[8]][1]

2*(1-pt(-0.431,34))
pt(-0.431,34)
0.669
1-pt(-0.431,34)
```
> i. Write E(y/x).

- E(SalesThisQuarter/Software = 1) = `r summary(softwareslr)$coefficients[1,1]`.

- E(SalesThisQuarter/Software = 2) = `r summary(softwareslr)$coefficients[1,1] + summary(softwareslr)$coefficients[2,1]`.

- E(SalesThisQuarter/Software = 3) = `r summary(softwareslr)$coefficients[1,1] + summary(softwareslr)$coefficients[3,1]`.

> ii. What proportion of the variation in sales this quarter is explained by software package? 

The proportion of the variation in sales this quarter that can be explained w the software package is `r summary(softwareslr)[[8]][1]`.

 > iii. What is the null hypothesis for testing whether software package has any effect on sales? 
 
 H0: The type of software package has no effect on sales this quarter. 

> v. Give the p-value. 

> vi. Do you reject H0 at α = 0.05? 

> vii. Are the results statistically significant at the 0.05 level? 

# Question 2 

> 2.Download the “Explosives dataset” from Moodle. Fit a simple linear regression, relating the deflection of 
galvonometer (Y) to the area of the wires on the coupling (X). Complete the following parts. 

```{r import2}
pigs <- as_tibble(read_table("Datasets/pig.txt",
col_names = TRUE
))
```

