---
title: "YuHW01ST430"
author: "Haozhe (Jerry) Yu"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
library(ggplot2)
library(dplyr)
library(tinytex)
library(reticulate)
```

# Question 1

>1. A regression analysis relating test scores (Y) to training hours (X)  produced the following fitted equation: ğ‘¦Ì‚ = 25 âˆ’ 0.5ğ‘¥.


>a. What is the fitted value of the response variable corresponding to  x = 7?

  25 - 0.5(7) = **3.5**
  
> b. What is the residual corresponding to the data point with x = 3 and  y = 30? Is the point above or below the line? Why?

  30-[25-(0.5*3)]= 6.5. The point is above the regression line because 30 > 23.5. 
  
> c. If x increases 3 units, how does ğ‘¦Ì‚ change?

  yÌ‚ decreases by 1.5.
  
d. An additional test score is to be obtained for a new observation at 
x = 6. Would the test score for the new observation necessarily be 22? 
Explain.
e. The error sums of squares (SSE) for this model was found to be 7. If 
there were n = 16 observations, provide the best estimate for ğœ
2
.


```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
