---
title: "YuHW00ST430"
author: "Haozhe (Jerry) Yu"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(tinytex)
```

# Question 1

>  The dataset sat.txt comes from a study entitled "Getting What You Pay For: The Debate 
Over Equity in Public School Expenditures.'' 
Make a numerical and graphical summary of the data, commenting on any features that you find interesting. 
Make sure you write something about anything you choose to present. You are not expected to make any 
substantive conclusions from the data.
Do some short numerical summaries of the data, commenting on any features that you 
find interesting.
(Please follow the R Codes given in Prostate Data and Pima Data)

First we input the SAT data from a local txt file. 


```{r import}
sat <-  as_tibble (read.delim2(
  "Datasets/SAT.txt",
  header = TRUE,
  sep = "",
  col.names = c("State", as.character(
    read.delim2(
      "Datasets/SAT.txt",
      nrows = 1,
      sep = "",
      header = FALSE
    )
  ))
))
sat$expend <- 1000 * as.numeric(sat$expend)
sat$salary <- 1000 * as.numeric(sat$salary)
sat$ratio <- as.numeric(sat$ratio)
sat
```

Then we run Pima and Prostate Style preanalysis.

```{r analysis}
summary(sat)

```


## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
