---
title: "Hw03ST430Yu"
author: "Haozhe (Jerry) Yu"
date: "`r Sys.Date()`"
output: 
  pdf_document:
    latex_engine: xelatex
---

```{r setup, include=FALSE}
library(ggplot2)
library(ggpmisc)
library(dplyr)
library(tinytex)
library(usethis)
library(lmtest)
library(car)
```

# Question 1

```{r import 1}
educ <- as_tibble(read.table("https://users.stat.ufl.edu/~rrandles/sta4210/Rclassnotes/data/textdatasets/KutnerData/Chapter%20%201%20Data%20Sets/CH01PR28.txt", 
            #sep = "", 
            strip.white=TRUE,
            col.name = c("Crime.Rate","High.School.Diploma")
            ))
educ
```


## a. Find the least squares regression equation to predict  the crime rate from the percent of individuals having at least a high school education. [Paste R or SAS output and then answer your question]

```{r model}
educm <- lm(Crime.Rate~High.School.Diploma,data=educ)

str(educm)
```

The equation to predict crime rate (per 100,000 residents) from the percent of individuals in a country with at least a high school diploma is 

Crime Rate = `r educm$coefficient[1]` + `r educm$coefficient[2]`High School Percent

## b. Give the ANOVA Table for this regression analysis. [Paste R or SAS output]

## c. Find SSE and MSE for this model.

## d. What is the estimate of σ from this analysis?

## e. What percent of the variation in crime rates can be explained by the percent of high school graduates?

## f. What is the correlation between crime rates and percent of high school graduates?

## g. Based on your ANOVA table, is the linear relationship between X and Y statistically significant? Be sure to give an appropriate null and alternate hypothesis, test statistic, its associated degrees of freedom, and the p-value.

## h. Give a scatter plot of crime rates vs. percent of high school graduates, with the regression line. Comment about linearity 

## i. Give the Residual Plot (residuals vs. fitted values). Test for Non-Linear and Non-constant variance.

## j. Conduct Breusch-Pagan Test for the constancy of the error variance. Be sure to give an appropriate null and alternate hypothesis, test statistic, its associated degrees of freedom, and the p-value.

## k. Index Plot to test for Independence of errors.

## l. Conduct Durbin-Watson Test. Be sure to give an appropriate null and alternate hypothesis, test statistic and the p-value.

## m. Outlier deduction test [Plot standardized Residuals versus fitted values]

## n. Give a Histogram of the residuals and the density curve. Comment about the distribution of residuals.


## o. Give a QQ-plot of the residuals to test for normality of error terms. Comment about the distribution of residuals.

## p. Conduct a Shapiro-Wilk Test on the residuals. Be sure to give an appropriate null and alternate hypothesis, test statistic and the p-value. Give the p-value for this test and explain what this means in terms of our model assumptions.

# Question 2

## a. Give a scatter plot 

## b. Find the least squares regression.

## c. Give the Residual Plot (residuals vs. fitted values). Test for Non-Linear and Non-constant variance.

## d. Conduct Breusch-Pagan Test for the constancy of the error variance.

## e. Index Plot to test for Independence of errors.

## f. Conduct Durbin-Watson Test.

## g. outlier deduction test. [Plot standardized Residuals versus fitted values]

## h. Give a Histogram of the residuals.

## i. Give a QQ-plot of the residuals. Normality of error terms.

## j. Conduct a Shapiro-Wilk Test on the residuals. Give the p-value for this test and explain what this means in terms of our model assumptions.

## k. Give the ANOVA Table for this regression analysis. Based on your ANOVA table, is the linear relationship between X and Y statistically significant? Be sure to give an appropriate test statistic, its associated degrees of freedom, and the p-value.


# Question 3

## a. Based on your ANOVA table, is the linear relationship between X and Y statistically significant? Be sure to give an appropriate null and alternate hypothesis, test statistic, its associated degrees of freedom, and the p-value.

## b. Give a scatter plot of clot vs. sur_time, with the 
regression line. Comment about linearity 

## c. Give the Residual Plot (residuals vs. fitted values). Test for Non-Linear and Non-constant variance.

## d. Conduct Breusch-Pagan Test for the constancy of the error variance. Be sure to give an appropriate null and alternate hypothesis, test statistic, its associated degrees of freedom, and the p-value.

## e. Index Plot to test for Independence of errors.

## f. Conduct Durbin-Watson Test. Be sure to give an appropriate null and alternate hypothesis, test statistic and the p-value.

## g. Outlier deduction test [Plot standardized Residuals versus fitted values]

## h. Give a Histogram of the residuals and the density curve. Comment about the distribution of residuals.

## i. Give a QQ-plot of the residuals to test for normality of error terms. Comment about the distribution of residuals.

## j. Conduct a Shapiro-Wilk Test on the residuals. Be sure to give an appropriate null and alternate hypothesis, test statistic and the p-value. Give the p-value for this test and explain what this means in terms of our model assumptions